---
format:
  gfm:
    output-file: README.md
editor: source
---

```{r}
#| include: false

knitr::opts_chunk$set(
  fig.align = "center",
  collapse = TRUE,
  comment = "#>"
)
```

# inti

[![CRAN](https://img.shields.io/cran/v/inti)](https://cran.r-project.org/package=inti)
[![DOI](https://zenodo.org/badge/82401374.svg)](https://zenodo.org/badge/latestdoi/82401374)
[![R-CMD-check](https://github.com/Flavjack/inti/workflows/R-CMD-check/badge.svg)](https://github.com/Flavjack/inti/actions)
[![Downloads](https://cranlogs.r-pkg.org/badges/grand-total/inti?color=brightgreen)](https://r-pkg.org/pkg/inti)

The **inti** package is part of the **inkaverse** project, which develops tools and workflows for plant science, experimental design, data analysis, and scientific writing. The package supports researchers during experiment planning and field data collection through **Tarpuy**, statistical analysis and visualization through **Yupana**, and the preparation of scientific outputs.

More information about the project is available at <https://inkaverse.com/>.

## Installation

The stable version of **inti** can be installed from CRAN:

```{r}
#| eval: false

install.packages("inti")
```

To install the latest development version directly from GitHub, it is recommended to use **pak**:

```{r}
#| eval: false

if (!requireNamespace("pak", quietly = TRUE)) {
  install.packages("pak")
}

pak::pkg_install("flavjack/inti")
```

To install a specific version from CRAN (e.g., version 0.4.4):

```{r}
#| eval: false

if (!requireNamespace("pak", quietly = TRUE)) {
  install.packages("pak")
}

pak::pkg_install("inti@0.4.4")
```

After installation, load the package:

```{r}
#| eval: false

library(inti)
```

## Shiny Applications

The **inti** package includes two Shiny applications that provide graphical interfaces for experimental design, data management, statistical analysis, and result visualization.

### Installing Application Dependencies

The applications can be accessed directly from the **RStudio Addins** menu or launched from the R console.

**Yupana** is an interactive environment for statistical analysis and visualization of experimental data.

```{r}
#| eval: false

inti:::yupana_full()
```

**Tarpuy** is an interactive platform for experimental planning, field-book generation, treatment organization, and data management.

```{r}
#| eval: false

inti:::tarpuy_full()
```