---
title: "Single-Factor Design: RCBD"
description: > 
  Randomized Complete Block Design
vignette: >
  %\VignetteIndexEntry{DoE-1_RCBD}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{quarto::html}
knitr:
  opts_chunk: 
    collapse: true
    comment: '#>'
    echo: false
    warning: false
    message: false
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
source("https://raw.githubusercontent.com/Flavjack/inti/master/pkgdown/favicon/docs.r")
```

Planning an experiment follows a reproducible routine:

1. **Load required libraries:** Load `inti`, `knitr`, and `dplyr` packages.
1. **Define factor levels:** Set up lists with genotypes, treatments, and management factors.
1. **Dispatch design generator:** Choose between CRD, RCBD, Split-plot, or Augmented designs.
1. **Plot the field sketch:** Verify spatial layouts and serpentine/zigzag sequences.
1. **Label design:** Design the experimental labels to facilitate the data collection.
5. **Export to Field Book app:** Generate field-ready sheets with trait parameters.

```{r, echo=TRUE}
# Install packages and dependencies

library(inti)
library(tidyverse)
library(huito)
```

## Randomized Complete Block Design (RCBD)

The Randomized Complete Block Design is recommended when an environmental gradient is present in the field, grouping experimental units into homogeneous blocks to control spatial variability.

```{r, echo=TRUE}
# 1. Define factors: Bean genotypes and fertilization doses
factors_rcbd <- list(
  Genotype = c("Bean_01", "Bean_02", "Bean_03"),
  Fertilization = c("0", "50", "100")
)

# 2. Generate RCBD layout (3 genotypes x 3 doses = 9 treatments, 4 blocks = 36 )
rcbd_exp <- design_repblock(
  nfactors = 2,
  factors = factors_rcbd,
  type = "rcbd",
  rep = 4,
  zigzag = TRUE,
  seed = 2026
)

# Fieldbook preview
rcbd_exp$fieldbook %>% 
  head(10) %>% 
  knitr::kable(caption = "RCBD Fieldbook preview")

# Field layout visualization
tarpuy_plotdesign(
  data = rcbd_exp,
  factor = "Genotype",
  fill = c("plots", "Fertilization")
)
```

