---
title: "Two-Factors Design: CRD"
description: > 
  Factorial Complete Randomize Design
vignette: >
  %\VignetteIndexEntry{DoE-2_CRD}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{quarto::html}
knitr:
  opts_chunk: 
    collapse: true
    comment: '#>'
    echo: false
    warning: false
    message: false
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
source("https://raw.githubusercontent.com/Flavjack/inti/master/pkgdown/favicon/docs.r")
```

Planning an experiment follows a reproducible routine:

1. **Load required libraries:** Load `inti`, `knitr`, and `dplyr` packages.
1. **Define factor levels:** Set up lists with genotypes, treatments, and management factors.
1. **Dispatch design generator:** Choose between CRD, RCBD, Split-plot, or Augmented designs.
1. **Plot the field sketch:** Verify spatial layouts and serpentine/zigzag sequences.
1. **Label design:** Design the experimental labels to facilitate the data collection.
5. **Export to Field Book app:** Generate field-ready sheets with trait parameters.

```{r, echo=TRUE}
# Install packages and dependencies

library(inti)
library(tidyverse)
library(huito)
```

# Designs with Two Factors

When evaluating two factors, four designs become available: **CRD**, **RCBD**, **Split-plot RCBD**, and **Augmented**.

# Factorial Completely Randomized Design (Factorial CRD)

Recommended for multi-factor experiments under homogeneous conditions, such as temperature- and salinity-controlled germination assays in growth chambers.

```{r, echo=TRUE}
# 1. Define factors: Salinity levels and incubation temperatures
factors_crd_2f <- list(
  NaCl = c("0", "50", "100"),
  Temp = c("20", "25")
)

# 2. Generate factorial CRD layout
crd_exp_2f <- design_repblock(
  nfactors = 2,
  factors = factors_crd_2f,
  type = "crd",
  rep = 4,
  zigzag = TRUE,
  seed = 2026
)

# Fieldbook preview
crd_exp_2f$fieldbook %>%
  head(10) %>%
  knitr::kable(caption = "Factorial CRD Fieldbook preview")

# Spatial layout visualization
tarpuy_plotdesign(
  data = crd_exp_2f,
  factor = "NaCl",
  fill = c("plots", "Temp")
)
```
