---
title: "Single-Factor Design: CRD"
description: >
  Complete Randomize Design
vignette: >
  %\VignetteIndexEntry{DoE-1_CRD}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{quarto::html}
knitr:
  opts_chunk: 
    collapse: true
    comment: '#>'
    echo: false
    warning: false
    message: false
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
source("https://raw.githubusercontent.com/Flavjack/inti/master/pkgdown/favicon/docs.r")
```

Planning an experiment follows a reproducible routine:

1. **Load required libraries:** Load `inti`, `knitr`, and `dplyr` packages.
1. **Define factor levels:** Set up lists with genotypes, treatments, and management factors.
1. **Dispatch design generator:** Choose between CRD, RCBD, Split-plot, or Augmented designs.
1. **Plot the field sketch:** Verify spatial layouts and serpentine/zigzag sequences.
1. **Label design:** Design the experimental labels to facilitate the data collection.
5. **Export to Field Book app:** Generate field-ready sheets with trait parameters.

```{r, echo=TRUE}
# Install packages and dependencies

library(inti)
library(tidyverse)
library(huito)
```

# Completely Randomized Design (CRD)

The Completely Randomized Design is recommended when experimental units are homogeneous, such as germination chambers, lab assays, or controlled greenhouse benches.

```{r, echo=TRUE}
# 1. Define salinity levels (NaCl concentrations in mM)
factors_crd <- list(
  NaCl= c("0", "50", "100", "150", "200")
)

# 2. Generate CRD layout (5 treatments x 4 replications = 20 petri dishes/units)
crd_exp <- design_repblock(
  factors = factors_crd,
  type = "crd",
  rep = 4,
  zigzag = TRUE,
  seed = 2026
)

# Fieldbook preview
crd_exp$fieldbook %>% 
  head(10) %>% 
  knitr::kable(caption = "CRD Fieldbook preview")

# Layout on germination chamber shelves

tarpuy_plotdesign(
  data = crd_exp,
  factor = "NaCl",
  fill = c("plots", "NaCl")
)
```

