---
title: "Two-Factors Design: Split-Plot in RCBD"
description: > 
  Split-Plot in RCBD
vignette: >
  %\VignetteIndexEntry{DoE-2_SPLIT}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{quarto::html}
knitr:
  opts_chunk: 
    collapse: true
    comment: '#>'
    echo: false
    warning: false
    message: false
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
source("https://raw.githubusercontent.com/Flavjack/inti/master/pkgdown/favicon/docs.r")
```

Planning an experiment follows a reproducible routine:

1. **Load required libraries:** Load `inti`, `knitr`, and `dplyr` packages.
1. **Define factor levels:** Set up lists with genotypes, treatments, and management factors.
1. **Dispatch design generator:** Choose between CRD, RCBD, Split-plot, or Augmented designs.
1. **Plot the field sketch:** Verify spatial layouts and serpentine/zigzag sequences.
1. **Label design:** Design the experimental labels to facilitate the data collection.
5. **Export to Field Book app:** Generate field-ready sheets with trait parameters.

```{r, echo=TRUE}
# Install packages and dependencies

library(inti)
library(tidyverse)
library(huito)
```

# Designs with Two Factors

When evaluating two or more factors, four designs become available: **CRD**, **RCBD**, **Split-plot RCBD**, and **Augmented**.

## Split-Plot Design in RCBD

The Split-plot Design is recommended when one factor requires larger experimental units due to management constraints (such as irrigation) assigned to main plots, while a second factor (such as commercial quinoa varieties) is assigned to sub-plots within each main plot.

```{r, echo=TRUE}
# 1. Define factors: Irrigation regimes (main plots) and commercial quinoa varieties (sub-plots)
factors_split <- list(
  Irrigation = c("Full", "Deficit"),
  Variety    = c("Var_1", "Var_2", "Var_3")
)

# 2. Generate Split-plot layout: 2 main levels x 3 sub levels x 4 blocks = 24 plots
split_exp <- design_split(
  factors = factors_split,
  type = "split_rcbd",
  rep = 4,
  zigzag = TRUE,
  seed = 2026
)

# Fieldbook preview
split_exp$fieldbook %>% 
  head(10) %>% 
  knitr::kable(caption = "Split-plot Fieldbook preview")

# Field layout visualization
tarpuy_plotdesign(
  data = split_exp,
  factor = "Irrigation",
  fill = c("plots", "Variety")
)
```

