---
# title: "Manuscript Analysis"
format:
  html:
    toc: true
    toc-expand: true
    toc-depth: 4
    toc-location: left
    number-sections: true
    self-contained: true
    code-fold: false
editor_options: 
  chunk_output_type: console
execute: 
  warning: false
  message: false
  echo: true
---

# Project Setup

```{r}
#| label: setup

source('https://inkaverse.com/setup.r')
```

# Data import

```{r}
gs <- "url" %>% 
  as_sheets_id()

fb <- gs %>% 
  range_read(ss = .,sheet = "fb") %>% 
  mutate(across(c(1:4), ~ as.factor(.))) 

fb %>% str()
```

# Research Objectives

> https://docs.google.com/spreadsheets/d/1u6xgTT8jDFkHSMkK1iRrNVR-1iOINQidbEiaDCu2Nqw/edit?gid=504037395#gid=504037395

## RO 1

```{r}

```

## RO 2

```{r}

```

# Multivariate

```{r}
#| eval: false

library(factoextra)

dtx <- fb %>% 
  dplyr::select(!c("var1", "var2")) %>% 
  unite("rowname", c("varX", "varY"), remove = F, sep = "-") %>% 
  column_to_rownames() %>% 
  as.data.frame()

str(dtx)
  
mv <- dtx %>% 
  PCA(scale.unit = T
      , graph = F
      , quali.sup = c(1, 2))

var <- mv %>% 
  plot.PCA(choix = c("var"), cex=0.7)

cos <- mv %>% 
  fviz_pca_var(col.var = "cos2"
               , gradient.cols = inti:::paleta()
               , repel = TRUE
               , labelsize = 2)

ind <- mv %>% 
  plot.PCA(choix = c("ind")
           , cex = 0.7
           , label = "ind"
           , invisible = "quali"
           , col.hab = inti:::paleta()
           , habillage = "none"
           )

pcainfo <- factoextra::get_pca(mv); pcainfo

pcainfo$cor %>% kable(caption = "Variable correlations")
pcainfo$contrib %>% kable(caption = "Variable Contribution")
pcainfo$cos2 %>% kable(caption = "Variable Cos2")

cls <- mv %>% HCPC(nb.clust = -1, graph = FALSE)

map <- \() {plot.HCPC(cls, choice = "map", draw.tree = FALSE)}

cls$call$t$tree
cls$data.clust %>% 
  kable(caption = "Cluster membership of each individual")  
cls$desc.var$quanti %>% 
  kable(caption = "Quantitative variables characterizing each cluster")
cls$desc.var$category%>% 
  kable(caption = "Categorical variables significantly by cluster")  
table(cls$data.clust$clust) %>% 
  kable(caption = "Number of individuals assigned to each cluster")

plot <- list(var, cos, ind, map) %>% 
  plot_grid(plotlist = ., ncol = 2, labels = "auto") %>% 
  ggsave(plot = ., "manuscript/multivariate.jpg", units = "cm"
                , width = 30, height = 30)

plot %>% include_graphics()
```

## PCA information

```{r}
c1 <- fviz_eig(mv, 
                addlabels = TRUE,
                hjust = 0.05,
                barfill="white",
                barcolor ="darkblue",
                linecolor ="red") + 
  # ylim(0, 30) +
  labs(
    title = "PCA - percentage of explained variances",
    y = "Variance (%)") +
  theme_minimal()  +
  theme(
    axis.text.x = element_text(size = 12)
    , axis.text.y = element_text(size = 12)
  )

c1

c2 <- fviz_contrib(mv,
                     choice = "var", 
                     axes = 1, 
                     top = 30,
                     fill="white",
                     color ="darkblue"
                   , sort.val = "asc"
                   ) +
  # ylim(0, 8) +
  labs(title = "Dim 1 - variables contribution", x = "") +
  coord_flip()  +
  theme(
    axis.text.x = element_text(size = 12)
    , axis.text.y = element_text(size = 12)
  )

c2

c3 <- fviz_contrib(mv,
                     choice = "var", 
                     axes = 2, 
                     top = 30,
                     fill="white",
                     color ="darkblue"
                   , sort.val = "asc"
                   ) +
  # ylim(0, 8) +
  labs(title = "Dim 2 - variables contribution", x = "") +
  coord_flip()  +
  theme(
    axis.text.x = element_text(size = 12)
    , axis.text.y = element_text(size = 12)
  )

c3

lpanel <- c(c1, c2, c3) %>% 
  plot_grid(plotlist = ., ncol = 1, labels = "auto"
            , rel_heights = c(1, 1.2, 1.2))
  
rpanel <- \() {
  
pcainfo$cos2[, 1:2] %>% 
  corrplot(
    is.corr = F
    , method = "number",
    type = "full",
    tl.col = "black",
    tl.srt = 45,
    tl.cex = 0.9, 
    number.cex = 0.5, 
    number.digits = 1,
    cl.cex = 0.7,
    cl.pos = "r",
    cl.align = "r",
    cl.ratio = 4 
  )

}

grid <- c(lpanel, rpanel) %>% 
  plot_grid(plotlist = ., ncol = 2, labels = c("", "d"))

plot <- grid %>% 
  ggsave(plot = ., "manuscript/contribution.jpg"
         , units = "cm"
         , width = 28, height = 36)

plot %>% include_graphics()
```


# References

Broman, K. W., & Woo, K. H. (2017). Data organization in spreadsheets. The American Statistician, 72(1), 2–10. https://doi.org/10.1080/00031305.2017.1375989

Zuur, A. F., Ieno, E. N., & Elphick, C. S. (2009). A protocol for data exploration to avoid common statistical problems. Methods in Ecology and Evolution, 1(1), 3–14. https://doi.org/10.1111/j.2041-210x.2009.00001.x

Francoishusson. (2017, July 13). PCA course using FactoMineR | R-bloggers. R-bloggers. https://www.r-bloggers.com/2017/07/pca-course-using-factominer/

Kozak, M., & Piepho, H. (2017). What’s normal anyway? Residual plots are more telling than significance tests when checking ANOVA assumptions. Journal of Agronomy and Crop Science, 204(1), 86–98. https://doi.org/10.1111/jac.12220

Tanaka, E., & Hui, F. K. C. (2019). Symbolic formulae for linear mixed models. In Communications in computer and information science (pp. 3–21). https://doi.org/10.1007/978-981-15-1960-4_1

Schielzeth, H., Dingemanse, N. J., Nakagawa, S., Westneat, D. F., Allegue, H., Teplitsky, C., Réale, D., Dochtermann, N. A., Garamszegi, L. Z., & Araya‐Ajoy, Y. G. (2020). Robustness of linear mixed‐effects models to violations of distributional assumptions. Methods in Ecology and Evolution, 11(9), 1141–1152. https://doi.org/10.1111/2041-210x.13434

Jambor, H. K. (2025). A checklist for designing and improving the visualization of scientific data. Nature Cell Biology, 27(6), 879–883. https://doi.org/10.1038/s41556-025-01684-z
