---
title: "Stagewise mixed-model analysis"
description: 'Comparison: asreml and H2cal'
output: rmarkdown::html_vignette
bibliography: [../files/pkgs.bib, ../files/book.bib]
link-citations: true
colorlinks: yes
csl: https://www.zotero.org/styles/crop-science
vignette: >
  %\VignetteIndexEntry{Stagewise mixed-model analysis}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
source("https://raw.githubusercontent.com/Flavjack/inti/master/pkgdown/favicon/docs.r")
knitr::opts_chunk$set(echo = TRUE)
```

Best linear unbiased prediction (BLUP) is used in plant breeding experiments for the selection of the best lines in multi-environmental trials (MET).

In this section, a comparison will be made between the `H2Cal()` function of the inti package [@R-inti] and the code using the `asrml` package published by @buntaran2020Cross. 

For the comparison we will use the two methods described by @buntaran2020Cross. (i) The "two-stage" analysis, using two stages, the first to calculate the BLUEs and the second the BLUPs. (ii) The "one stage" analysis using only one model to calculate the BLUPs.

# Load data

```{r}
library(inti)
library(purrr)
library(dplyr)
fb <- inti::met
```

# Two-stage analysis

## asreml

```{r, results='hide', cache=TRUE}
library(asreml)
library(data.table)
library(plyr)
library(stringr)
asreml.options(maxit=100) # Set asreml iteration

############################
##### Stage I LSMEANS #####
##### per location   #####

ww <- data.table(fb)

##### Make column Zone_Loc #####

trials     <- nlevels(ww$env)
envs <- levels(ww$env)

##### Make data list for Stage I #####
data_list <- matrix(data=list(), nrow=length(envs), ncol=1, 
                    dimnames=list(envs, c("data_Set")))

##### Make a list of Trials #####
for(i in 1:trials){
  print(i)
  b <- levels(ww$env)
  c <- b[i]
  env <- as.factor(c)
  env <- data.table(env)
  f <- merge(ww,env,by="env")
  assign(paste0("data_", b[i]), f)
  data_list[[i, "data_Set" ]] <- f
  
  rm(b, c, f, env)
}

data_list <- data.table(ldply(data_list[, "data_Set"], data.frame, .id="env"))

stgI_list <- matrix(data=list(), nrow=length(envs), ncol=1, 
                    dimnames=list(envs, c("lsmeans")))

asreml.options(maxit=100) # Set asreml iteration

############################
##### Stage I LSMEANS #####
##### per location   #####

for (i in envs){
  
  edat <- droplevels(subset(ww, env == i))
  
  print(i)
  
  mod.1 <- asreml(fixed     = yield ~ cultivar,
                  random      = ~ rep + rep:alpha,
                  data        = edat,
                  predict     = predict.asreml(classify = "cultivar"))
  
  update.asreml(mod.1)
  print(summary.asreml(mod.1)$varcomp)
  
  blue <- predict(mod.1, classify="cultivar", levels=levels(edat$cultivar), vcov=TRUE,aliased = T) # get the lsmeans
  blue.1 <- data.table(blue$pvals)[, c(1:3)] 
  names(blue.1) <- c("cultivar", "yield_lsm", "se")
  blue.1[ , ':='(var=se^2, smith.w=diag(solve(blue$vcov)))] # calculate the Smith's weight
  
  stgI_list[[i, "lsmeans" ]] <- blue.1 # put all the results of Stage 1 in the list
  
  rm(Edat,mod.1, blue, blue.1)
}

#######################################################
##### Preparing dataset of Stage I for Stage II ######

##### Unlist the results of Stage I and format as data.table #####
stgII_list <- data.table(plyr::ldply(stgI_list[, "lsmeans"], data.frame, .id="env"))

stgII_list$zone<- factor(str_split_fixed(stgII_list$env, "_", 2)[,1]) # Make Zone column by split the record in Zone_Loc column
stgII_list$location <- factor(str_split_fixed(stgII_list$env, "_", 3)[,2])  # Make Location by split the record in Zone_Loc column
stgII_list$year <- factor(str_split_fixed(stgII_list$env, "_", 3)[,3]) # Make Year by split the record in Zone_Loc column

blues.asreml <- stgII_list

############################
##### Stage II BLUPs ######
##### Zone analysis #####

model <- asreml(yield_lsm  ~ zone,
                random    = ~cultivar + zone:location + zone:cultivar + cultivar:zone:location,
                weights   = smith.w,
                family    = asr_gaussian(dispersion=1.0), # fix residual variance to 1
                data      = blues.asreml,
                predict   = predict.asreml(classify = "cultivar")
                )

update.asreml(model)

# print(summary.asreml(model)$varcomp) # print the variance components

blups.asrml <- data.frame((model$predictions$pvals[1:4]))
```

## H2cal

```{r}
library(inti)
library(purrr)

#> First stage

envs <- levels(fb$env)

model <- 1:length(envs) %>% map(function(x) {
  
  model <- fb %>% filter(env %in% envs[x]) %>% 
    
    H2cal(trait = "yield"
          , gen.name = "cultivar"
          , rep.n = 4
          , fixed.model = "0 + (1|rep) + (1|rep:alpha) + cultivar"
          , random.model = "1 + (1|rep) + (1|rep:alpha) + (1|cultivar)"
          # , plot_diag = T
          , emmeans = F
          )
  
  blues <- model$blues %>% mutate(trial = levels(fb$env)[x])
  
  })

blues.h2cal <- bind_rows(model) %>% 
  separate(trial, c("zone", "location", "year")) %>% 
  mutate(across(c(yield, smith.w), as.numeric)) %>% 
  mutate(across(!c(yield, smith.w), as.factor))

#> Second stage

met <- blues.h2cal %>%
  mutate(across(!yield, as.factor)) %>%
  H2cal(trait = "yield"
        , gen.name = "cultivar"
        , rep.n = 4
        , loc.n = 18
        , loc.name = "location"
        , fixed.model = "0 + zone + (1|zone:location) + (1|zone:cultivar) + cultivar"
        , random.model = "1 + zone + (1|zone:location) + (1|zone:cultivar) + (1|cultivar)"
        # , plot_diag = T
        , emmeans = T
        # , weights = blues.h2cal$smith.w
        )

blups.h2cal <- met$blups
```

- First stage
  - Fixed model with `0 +` for avoid intercep and calculate all the BLUEs.
  - `emmeans = F` to calculate the Smith weitghts in the first stage

## BLUEs comparison

```{r}
blues.comp <- merge(blues.asreml 
                    , blues.h2cal 
                    , by = c("cultivar", "zone", "location"))

# plot(blues.comp$yield, blues.comp$yield_lsm)
rs <- cor(blues.comp$yield, blues.comp$yield_lsm) 
cat("r =", rs)

blues.comp %>% web_table(digits = 4)
```

The BLUEs correlation between `H2Cal` and `asrml` is (r = `r rs`)

## BLUPs comparison

```{r}
blups.comp <- merge(blups.asrml, blups.h2cal
                    , by = c("cultivar"))

# plot(blups.comp$yield, blups.comp$predicted.value)
rs <- cor(blups.comp$yield, blups.comp$predicted.value) 
cat("r =", rs)

blups.comp %>% web_table(digits = 4)
```

The BLUPs correlation between `H2Cal` and `asrml` is (r = `r rs`)

# Single-stage analysis

## asreml

```{r, results='hide', cache=TRUE}
library(asreml)

options("scipen"=100,"digits"= 4 )
asreml.options(maxit=100) # Set asreml iteration

##### Fit a single-stage model #####
## incomplete block and replicate location-specific
## location-specifice residual variance

mod <- asreml(fixed       = yield ~ zone,
              random      = ~ rep:at(location) + rep:alpha:at(location) + zone:location + 
                             cultivar + cultivar:zone:location+ cultivar:zone,
              residual    = ~ dsum(~(units)|location),
              data        = fb,
              predict     = predict.asreml(classify = "cultivar"))

update.asreml(mod)

blups.asreml <- data.frame((mod$predictions$pvals[1:4])) 
```

## H2cal

```{r}
library(inti)

model <- fb %>% 
  H2cal(trait = "yield"
        , gen.name = "cultivar"
        , loc.name = "location"
        , rep.n = 2
        , loc.n = 18
        , fixed.model = "0 + zone + (1|rep:location) + (1|rep:alpha:location) + (1|zone:location) + (1|cultivar:zone) + (1|cultivar:zone:location) + cultivar"
        , random.model = "1 + zone + (1|rep:location) + (1|rep:alpha:location) + (1|zone:location) + (1|cultivar:zone) + (1|cultivar:zone:location) + (1|cultivar)"
        , summary = T
        , emmeans = T
        # , plot_diag = T
        )

blups.h2cal <- model$blups 
```

## BLUPs comparison

```{r}
blups.comp <- merge(blups.asreml, blups.h2cal, by = c("cultivar"))

# plot(blups.comp$predicted.value, blups.comp$yield)
rs <- cor(blups.comp$predicted.value, blups.comp$yield) 
cat("r =", rs)

blups.comp %>% web_table(digits = 4)
```

The BLUPs correlation between `H2Cal` and `asrml` is (r = `r rs`)

# References

::: {#refs}
:::

